Mental simulation as Bayesian Quadrature

### 2013-05-29 man/nips-2013.tex
Include analysis tex file
Use analysis macros

@@

The gold standard always evaluates $S$ at all angles, so the
correlation between it and the true angle of rotation is
undefined. The mean error of responses was $\ME{}=0.01$, indicating
that $Z$ is a sufficient approximation of $\int_R p(X_b\ \vert\
X_R)p(R)\ \mathrm{d}R$.

@@

The correlation between the number of rotations computed by the
\naive{} model and the true angle of rotation was $\rho=-0.11$. The
shape of the data is more interesting (see Figure \ref{fig:rotations},
left panel): the \naive{} model actually corresponds quite well to
the true angle of rotation for $R<\frac{\pi}{2}$ ($\rho=0.64$). This
is unsurprising, because the closer the true angle is to zero, the
less the model has to rotate, and the less likely it will get stuck on
local maxima. Thus, it is more likely to locate the global maximum,
which generally corresponds to the true angle of rotation. For
$R>\frac{\pi}{2}$, we see an increasing tendency to under-rotate
($\rho=-0.42$); this is likely because it finds a local maxima and
ends prematurely.

The \naive{} model's response error rate was better than chance, but
still quite high, with $\ME{}=0.23$. Closer inspection reveals that
the bulk of this comes from $h_1$ stimuli ($\ME{}=0.36$
vs. $\ME{}=0.1$ for $h_0$ stimuli). This is, as above, probably
related to finding only local maxima: if the model finds a local
maxima which is low enough, the area under the estimated curve will
not be large enough to accept $h_1$.

This is reflected in the accuracy in estimating $Z$ as well
($\MSE{}=0.14$). Figure \ref{fig:accuracy} (left panel) shows
individual estimates for each stimulus, color-coded by true
hypothesis. The model underestimates $Z$ for significant portion of
the $h_1$ stimuli. On the other $h_1$ stimuli, the \naive{} model
actually tends to overestimate because linear interpolation does not
account for symmetry of global maxima (e.g., Figure \ref{fig:li}).

@@

Like the \naive{} model, the parametric model uses hill-climbing as a
sampling strategy, so they have identical correlations with the true
angle of rotation. However, given the same $\mathbf{R}$, the
parametric model estimates $Z$ differently, thus giving rise to
different response and estimated $Z$ error rates. The response error
for the parametric model was higher than the \naive{} model, with
$\ME{}=0.26$. Breaking this down, we see that the error for $h_1$
stimuli is $\ME{}=0.5$: in other words, the parametric model is at
chance when determining whether two identical shapes are the
same. This is because it never overestimates (see Figure
\ref{fig:accuracy}, center panel). So, it accurately estimates $Z$ for
stimuli which the \naive{} model would overestimate ($\MSE{}=0.07$),
but still has poor performance when $Z$ is underestimated. Indeed, if
we exclude stimuli for which the \naive{} model overestimates ($Z\geq
0.3$), the error for the \naive{} model lowers to match that of the
parametric model.

@@

The rotations computed by the nonparametric model were strongly
correlated with the true rotations ($\rho=0.73$, see Figure
\ref{fig:rotations}, right panel). This is because the nonparametric
model does not get stuck as easily on local optima: it will continue
rotating until it is confident that its estimate of $Z$ is accurate.

Correspondingly, the nonparametric model is much more accurate in
choosing the correct hypothesis ($\ME{}=0.06$). As with the \naive{}
and parametric models, it mostly errs on $h_1$ stimuli
($\ME{}=0.1$). This appears to mostly be the result of choosing the
incorrect initial direction of rotation.

Similarly, because the nonparametric model actively attempts to obtain
an accurate estimate of $Z$, its estimate of $Z$ is indeed fairly
accurate ($\MSE{}=0.10$). This is slightly higher than the
nonparametric model, but is due to a few outliers. If we again exclude
stimuli for which $Z$ is greatly overestimated ($Z\geq 0.3$), the
nonparametric model's error decreases significantly ($\MSE{}=0.03$).

>>>>

\include{nips-2013-analyses}

@@

The gold standard always evaluates $S$ at all angles, so the
correlation between it and the true angle of rotation is
undefined. The mean error of responses was $\GSME{}$, indicating that
$Z$ is a sufficient approximation of $\int_R p(X_b\ \vert\ X_R)p(R)\
\mathrm{d}R$.

@@

The correlation between the number of rotations computed by the
\naive{} model and the true angle of rotation was $\Naivecorr{}$. The
shape of the data is more interesting (see Figure \ref{fig:rotations},
left panel): the \naive{} model actually corresponds quite well to the
true angle of rotation for $R<\frac{\pi}{2}$ ($\Naivecorrsm{}$). This
is unsurprising, because the closer the true angle is to zero, the
less the model has to rotate, and the less likely it will get stuck on
local maxima. Thus, it is more likely to locate the global maximum,
which generally corresponds to the true angle of rotation. For
$R>\frac{\pi}{2}$, we see an increasing tendency to under-rotate
($\Naivecorrbg{}$); this is likely because it finds a local maxima and
ends prematurely.

The \naive{} model's response error rate was better than chance, but
still quite high, with $\NaiveME{}$. Closer inspection reveals that
the bulk of this comes from $h_1$ stimuli ($\NaiveMEsame{}$
vs. $\NaiveMEdiff{}$ for $h_0$ stimuli). This is, as above, probably
related to finding only local maxima: if the model finds a local
maxima which is low enough, the area under the estimated curve will
not be large enough to accept $h_1$.

This is reflected in the accuracy in estimating $Z$ as well
($\NaiveMSE{}$). Figure \ref{fig:accuracy} (left panel) shows
individual estimates for each stimulus, color-coded by true
hypothesis. The model underestimates $Z$ for significant portion of
the $h_1$ stimuli. On the other $h_1$ stimuli, the \naive{} model
actually tends to overestimate because linear interpolation does not
account for symmetry of global maxima (e.g., Figure \ref{fig:li}).

@@

Like the \naive{} model, the parametric model uses hill-climbing as a
sampling strategy, so they have identical correlations with the true
angle of rotation. However, given the same $\mathbf{R}$, the
parametric model estimates $Z$ differently, thus giving rise to
different response and estimated $Z$ error rates. The response error
for the parametric model was higher than the \naive{} model, with
$\VMME{}$. Breaking this down, we see that the error for $h_1$ stimuli
is $\VMMEsame{}$: in other words, the parametric model is at chance
when determining whether two identical shapes are the same. This is
because it never overestimates (see Figure \ref{fig:accuracy}, center
panel). So, it accurately estimates $Z$ for stimuli which the \naive{}
model would overestimate ($\VMMSE{}$), but still has poor performance
when $Z$ is underestimated. \TODO{Indeed, if we exclude stimuli for
  which the \naive{} model overestimates ($Z\geq 0.3$), the error for
  the \naive{} model lowers to match that of the parametric model.}

@@

The rotations computed by the nonparametric model were strongly
correlated with the true rotations ($\BQcorr$, see Figure
\ref{fig:rotations}, right panel). This is because the nonparametric
model does not get stuck as easily on local optima: it will continue
rotating until it is confident that its estimate of $Z$ is accurate.

Correspondingly, the nonparametric model is much more accurate in
choosing the correct hypothesis ($\BQME{}$). As with the \naive{} and
parametric models, it mostly errs on $h_1$ stimuli ($\BQMEsame$). This
appears to mostly be the result of choosing the incorrect initial
direction of rotation.

Similarly, because the nonparametric model actively attempts to obtain
an accurate estimate of $Z$, its estimate of $Z$ is indeed fairly
accurate ($\BQMSE{}$). This is slightly higher than the nonparametric
model, but is due to a few outliers. \TODO{If we again exclude stimuli
  for which $Z$ is greatly overestimated ($Z\geq 0.3$), the
  nonparametric model's error decreases significantly
  ($\MSE{}=0.03$).}
